\bf Concurrent Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes

### 2022-10-10 writing/draft.tex
Manuscript revised

	\if0\blind
	{
		\title{\bf Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes}
		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
		$^a$ Department of Industrial and Manufacturing Engineering, \\
		The Pennsylvania State University, State College, PA USA \\
        $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
        Wright State University, Dayton, OH USA}
		\date{}
		\maketitle
	} \fi

@@

\begin{abstract}
Potassium channels (K\textsubscript{v}) are responsible for repolarizing the cardiac action potential. Different K\textsubscript{v} isoforms contribute to various phases of repolarization. In the mouse ventricle, repolarization roughly consists of two kinetic components: the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1). A data assimilation method, combining experimental data with mathematical/computational models, is used to estimate information on K\textsubscript{v} isoforms, because only the sum of their activities can be measured in the form of currents (I\textsubscript{K}). Most existing studies separate K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure. However, this traditional approach has limitations to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell one-by-one individually. There is an urgent need to develop new methods that simultaneously model the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of pathological mechanisms. We propose a novel \textit{concurrent data assimilation} framework that utilizes biophysics-based computer models of K\textsubscript{v} to delineate gating kinetics in the multiple observations concurrently to cope with cellular-level dynamics. We apply the proposed framework to glyco-altered I\textsubscript{K} recordings from our previous studies. Experimental results show that the proposed method can discriminate characteristics between the control and experimental groups. This proposed method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
\end{abstract}

\noindent%
{\it Keywords:} Data assimilation, model calibration, potassium channels, kinetics modeling, glycosylation

@@

\section{Introduction}
Cardiac voltage-gated potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential (AP) in cardiomyocytes. They are indispensable in the electrical conduction system of the heart. Even modest changes in K\textsubscript{v} activities can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}. Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}. A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, have shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics,nagai2016aberrant}. Despite its importance, the underlying mechanism is still largely unknown how changes in glycosylation contribute to heart disease onset and progression. Our \textit{in-vitro} experiments showed that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}. In addition, we verified our experimental results via \textit{in-silico} studies and estimate how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}. \textbf{Figure~\ref{fig:framework_diagram}} illustrates this feedback loop of studying heart diseases, which will be referred to as the data assimilation framework that combines experimental data with mathematical/computational models.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/overall_diagram.pdf}
    \caption{Conceptual diagram to illustrate a data assimilation framework.}
    \label{fig:framework_diagram}
\end{figure}

However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies: K\textsubscript{v} have diverse isoforms, with each of them showing unique biophysical properties and playing different roles in the AP repolarization \citep{nerbonne2005molecular}. In contrast, there is a predominant isoform for Na\textsubscript{v} and Ca\textsubscript{v} \citep{abriel2010cardiac,benitah2010type}. Although it is highly desirable to observe the activities of each K\textsubscript{v} isoform separately, only collective outputs can be measured because of their slightly overlapping voltage-dependence of gating \citep{brouillette2004functional}. Specifically, only the sum of K\textsuperscript{+} currents (I\textsubscript{K}) can be recorded in whole-cell voltage-clamp experiments. Typically, a form of curve fitting is used to estimate individual K\textsuperscript{+} currents, in which parameters of the functional form of the currents are searched, thereby the sum of them best describing I\textsubscript{K} recordings \citep{brunet2004heterogeneous}. Then the shape parameters are used to calibrate computer models \citep{kim2022simulation}, or kinetics information is added to the data assimilation process with additional curve fitting if applicable \citep{du2017silico}. However, this traditional approach is limited in its ability to rigorously determine kinetics and gating of K\textsubscript{v} isoforms in two aspects: 1) This two-step curve-fitting method does not consider the dynamics of the underlying kinetics and current outputs by separating the two processes working together and simplifying them with curve functions, and 2) it does not account for the cellular-level dynamics according to changes in protocols, as one current trace is analyzed separately. These limitations not only hinder \textit{in-silico} modeling results but, in some cases, make it impossible if modeling relies on kinetic data estimated from experimental data \citep{kim2022simulation}. Therefore, there is an urgent need to develop new data assimilation methods and tools that delineate the kinetics of K\textsubscript{v} isoforms for better understanding their roles in heart diseases.

This paper presents a novel approach to data assimilation for analyzing whole-cell K\textsubscript{v} current traces (I\textsubscript{K}) and modeling the underlying kinetics of K\textsubscript{v} isoforms. A schematic flowchart of the proposed framework is shown in \textbf{Figure~\ref{fig:flow_chart}B} and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}A}. First, we designed biophysical computer models that simulate activities of ion channels to replace simple current-shape functions in curve fitting. Computer models of ion channels are a system of differential equations consisting of gating variables and kinetic functions that continuously simulate ion channel activities over time. Parameters in kinetic functions control the gating of ion channels and, in turn, determine current outputs. Second, a sensitivity analysis was performed to screen out parameters that marginally influence the model outputs. Third, we developed a model calibration that minimizes discrepancy between model outputs and experimental data generated from multiple protocols. Last, we made estimations and inferences of impacts of disease-related perturbations on potassium channel kinetics and used low-dimensional embedding to show cellular variability. Experimental results of the proposed method applied to our glycosylation data provide kinetics information and differences between the healthy control and the disease group.
\begin{figure}
    \centering
    \includegraphics{figs/flow_chart.pdf}
    \caption{Flowcharts of (A) the current approach and (B) the proposed method.}
    \label{fig:flow_chart}
\end{figure}

The main advantage of this approach is that it analyzes I\textsubscript{K} recordings based on the biophysical models with considering cellular level dynamics according to changes in protocols for studying heart diseases. Our contributions are summarized as follows:
\begin{itemize}
    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting methods.
    \item We developed a model calibration routine that harnesses multiple protocols simultaneously rather than a single protocol independently, resulting in comprehensive cellular-level modeling.
    \item The proposed approach builds separate K\textsubscript{v} models and calibrates them for each cell, which allows analyzing the cellular variability. 
    \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
\end{itemize}
The proposed data assimilation framework has strong potential to serve as a novel complementary data analytics tool to in-vitro experiments.

@@

\subsection{Data Assimilation and Model Calibration}
\textit{Data assimilation} is a method to find the optimal configuration and state of computational models by coupling them with experimental data. Experimental data $\mathcal{D}$ are observations of a real process $\mathcal{R}$ that represents scientific phenomena under investigation. The output of physical experiments $y^{\mathcal{D}}(x)$, given input $x$, inevitably contains errors for various reasons, such as noise in measurement or experimental environment. Suppose $\mathcal{D}$ and $\mathcal{R}$ can be related as follows in Equation~\ref{eq:rd_relation}, where $\epsilon$ is the error term. 
\begin{equation}
    \label{eq:rd_relation}
    y^{\mathcal{R}}(x) = y^{\mathcal{D}}(x) + \epsilon
\end{equation}
Let $y^{\mathcal{M}}(x|\theta)$ denote the output from a computer model $\mathcal{M}$, given parameters $\theta$. Assume that there are discrepancies $\delta(x|\theta)$ for the current states of parameters as follows in Equation~\ref{eq:dm_relation}. 
\begin{align}
    \label{eq:dm_relation}
    y^{\mathcal{D}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) \text{, so} \\
    y^{\mathcal{R}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) + \epsilon
\end{align}
Our goal in data assimilation is to calibrate $\theta$ to find the best model states that minimize $\delta(x|\theta)$, while satisfying biophysical constraints. By doing that, \textit{in-silico} models $\mathcal{M}$ are coupled with \textit{in-vitro} experimental data $\mathcal{M}$, which provides two complementary angles to study the real process $\mathcal{R}$.

In this study, $\delta$ is defined by the sum of root-mean-square errors (RMSEs) as in Equation~\ref{eq:rmse}, which measures deviations between experimental data and model predictions from the end of a holding potential $t_i^{(h)}$ to the end of voltage step $t_i^{(e)}$ across a range of protocols, $i=1,2,\cdots,n$.
\begin{equation}
    \label{eq:rmse}
    \delta = \sum_{i=1}^{n} \sqrt{\int_{t_i^{(h)}}^{t_i^{(e)}}\frac{(y_i^{\mathcal{D}}(t) - y_i^{\mathcal{M}}(t|\theta))^2}{t_i^{(e)}-t_i^{(h)}}dt}
\end{equation}
This approach differs from previous studies in two-fold. First, unlike the traditional curve fitting using the exponential model described in Equation~\ref{eq:expl_fitting}, the suggested method includes multiple protocols simultaneously for comprehensive cellular-level modeling. Second, it calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data \citep{du2015statistical, du2017silico, kim2022simulation}. Calibrating to the current traces themselves raises a challenge in optimization but has advantages when it is hard to estimate statistics from data reliably \citep{kim2022simulation}.

We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$. Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube. In this way, the optimization loop can be controlled by users, allowing them to blend their domain knowledge into the modeling. Note that it is possible because the suggested method is based on biophysical models that provide interpretability of $\theta$ in channel kinetics rather than the curve fitting. Global optimum is not guaranteed due to the complexity of our models. We developed the multi-random-start scheme to avoid local optima and find the solution as close to the optimum as possible. Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden. This work is implemented in MATLAB R2022a, and the code is available at xyz.

@@

\subsection{Kinetics Modeling} \label{s:results.kinetics}
The \textit{in-silico} modeling predicts that reduced N-glycosylations cause significant changes in channel kinetics and gating behavior. \textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results showing kinetics variables to voltage relationships of K\textsubscript{v}4.2. There is a slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.25 at voltages lower -10 mV (see \textbf{Figure~\ref{fig:kinetics_ikto}B}). Activation and inactivation time constants ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}. $\tau_{a}^{(1)}$ for MGAT1KO is greater than WT in range of $>$ -20 mV. $\tau_{i}^{(1)}$ for MGAT1KO shifted toward left. The shifts in SSA and SSI likely cause the reductions in I\textsubscript{Kto} magnitudes, and the increased time constants explain delays in the decaying phase observed in the \textit{in-vitro} experiments.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kinetics_ikto.pdf}
    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kto} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, and (D) inactivation time constant.}
    \label{fig:kinetics_ikto}
\end{figure}

\textbf{Figure~\ref{fig:kinetics_rec}} shows kinetics modeling results of the channels producing the delayed rectifier currents, I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, and steady-state current I\textsubscript{Kss}. The most prominent result is that time constants of activation and inactivation in I\textsubscript{Kslow1} ($\tau_{a}^{(2)}$ and $\tau_{i}^{(2)}$) are estimated to be significantly slow in MGAT1KO, i.e., $\sim$ 3.5 and 348.9 ms, respectively (see \textbf{Figure~\ref{fig:kinetics_rec}D} and \textbf{E}). Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO. SSI is also shifted $\sim$ 10 mV to the right. The slow time constants and increased SSI are likely to address the flattened and reduced currents of MGAT1KO. Time constant inactivation of I\textsubscript{Kslow2} ($\tau_{i}^{(3)}$) is also increased (see \textbf{Figure~\ref{fig:kinetics_rec}C}), but activation of I\textsubscript{Kss} ($\tau_{a}^{(4)}$) do not show significant differences between the two groups (see \textbf{Figure~\ref{fig:kinetics_rec}F}).
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kinetics_rec.pdf}
    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kslow1} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, (D) inactivation time constant, (E) I\textsubscript{Kslow2} inactivation time constant, and (F) I\textsubscript{Kss} activation time constant.}
    \label{fig:kinetics_rec}
\end{figure}

\subsection{Parameter Distributions Between WT and MGAT1KO} \label{s:results.distribution}
We estimated distributions of the calibrated kinetic parameters using histograms to analyze the impacts of reduced N-glycosylation further. \textbf{Figure~\ref{fig:hist_pkto}} shows the distributional differences of the kinetic parameters in the K\textsubscript{4.2} model between the two group. A notable fact about the figure is that MGAT1KO shows small variances for Parameters 1, 5, and 7 in transition $\alpha_{a}$ and $\beta_{a}$, which control kinetics activation of I\textsubscript{Kto}. Reduced glycosylation affects activation gating that plays a vital role in I\textsubscript{Kto}, as it is characterized by a high peak at the early depolarization and rapid inactivation. In addition, MGAT1KO causes a leftward shift in the distribution of the maximum conductance $G_{\mathrm{Kto}}$, resulting in smaller conduction than WT in general. It is likely due to the observed current reductions in MGAT1KO.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/hist_pkto.pdf}
    \caption{Distribution estimations of calibration parameters in the K\textsubscript{Kto} model.}
    \label{fig:hist_pkto}
\end{figure}

>>>>

	\if0\blind
	{
		\title{\bf Concurrent Data Assimilation for Modeling and Calibrating of Disease-altered Potassium Channel Kinetics in Mouse Cardiomyocytes}
		\author{Haedong Kim $^a$, Hui Yang $^a$, Andrew R. Ednie $^b$, and Eric S. Bennett $^b$ \\
		$^a$ Department of Industrial and Manufacturing Engineering, \\
		The Pennsylvania State University, University Park, PA USA \\
        $^b$ Department of Neuroscience, Cell Biology, and Physiology, \\
        Wright State University, Dayton, OH USA}
		\date{}
		\maketitle
	} \fi

@@

\begin{abstract}
Potassium channels (K\textsubscript{v}) are responsible for repolarizing the action potential in cardiomyocytes. Different K\textsubscript{v} isoforms and their currents contribute to various phases of repolarization. In the mouse ventricle, repolarization roughly consists of two kinetic components: the rapidly inactivating current I\textsubscript{Kto} (K\textsubscript{v}4.2), and the slowly inactivating currents I\textsubscript{Kslow1} and I\textsubscript{Kslow2} (K\textsubscript{v}1.5 and K\textsubscript{v}2.1). Data assimilation is used to estimate information on K\textsubscript{v} isoforms, because only the sum of their activities is measured in the form of currents (I\textsubscript{K}). Most existing studies separate K\textsuperscript{+} currents mathematically from I\textsubscript{K} recordings and make inference of kinetics by a curve-fitting decomposition procedure. However, this traditional approach has limitations to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting decomposition only relies on the shape of K\textsuperscript{+} currents without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell one-by-one individually. There is an urgent need to develop new methods that simultaneously model the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of pathological mechanisms. We propose a novel \textit{concurrent data assimilation} framework that utilizes biophysics-based computer models of K\textsubscript{v} to delineate gating kinetics in the multiple observations concurrently to cope with cellular-level dynamics. We apply the proposed method to glyco-altered I\textsubscript{K} recordings from our previous studies. Experimental results show that the proposed method can discriminate characteristics between the control and experimental groups. This proposed method has strong potential to pave a new way to analyze K\textsubscript{v} and heart diseases.
\end{abstract}

\noindent%
{\it Keywords:} Data assimilation, computational electrophysiology, potassium channels, kinetics modeling, glycosylation

@@

\section{Introduction}
Voltage-gated potassium channels (K\textsubscript{v}) are critical to the electrical conduction system of the heart. K\textsubscript{v} are responsible for repolarizing the cardiac action potential (AP), a change of membrane potential over time, representing the net electrical activity in a myocyte. The AP is produced by the orchestrated activities of ion channels. Voltage-gated sodium channels (Na\textsubscript{v}) initiates fast depolarization that drives the sharp upstroke of the AP. K\textsubscript{v} play antagonistic roles that repolarize the AP with the rapid inactivation of Na\textsubscript{v}. Repolarization consists of multiple phases, to which different K\textsubscript{v} isoforms contribute with fluxes of Ca\textsuperscript{2+} and Cl\textsuperscript{-} involved through designated ion channels. This complex process can be affected by even modest changes in K\textsubscript{v} activities. Aberrant repolarization can significantly affect the AP duration and the QT interval, which lead to fatal heart diseases \citep{ravens2008role}. Glycosylation is a co/posttranslational modification that is critical for protein functions including activities of K\textsubscript{v} and other voltage-gated ion channels (VGICs) \citep{ohtsubo2006glycosylation,ednie2012modulation}. A growing number of studies in diverse directions, ranging from genome-wise to proteomic/glycomic searches, have shown clear links between altered protein glycosylation and heart diseases \citep{yung2004gene,yang2015glycoproteins,miura2016glycomics,nagai2016aberrant}. Despite its importance, the underlying mechanism is still largely unknown how changes in glycosylation contribute to heart disease onset and progression. Our \textit{in-vitro} experiments showed that glycosylation regulations modulate VGIC activities and contribute to both electrical and contractile dysfunction \citep{ednie2013sialicNav1,ednie2015sialicKv,ednie2019reduced}. In addition, we verified our experimental results via \textit{in-silico} studies and estimate how changes in the lower level (VGICs) affect the functions in the higher level (cardiomyocytes), which is one of the powers of systematic computer models \citep{du2013silico,du2015statistical,du2017silico,kim2022simulation}. \textbf{Figure~\ref{fig:framework_diagram}} illustrates this feedback loop of studying heart diseases, which will be referred to as the data assimilation framework that combines experimental data with mathematical/computational models.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/overall_diagram.pdf}
    \caption{Conceptual diagram to illustrate a data assimilation framework.}
    \label{fig:framework_diagram}
\end{figure}

However, a distinctive feature of K\textsubscript{v} raises challenges in both \textit{in-vitro} and \textit{in-silico} studies: K\textsubscript{v} have various isoforms playing different roles in AP repolarization \citep{nerbonne2005molecular}, whereas dominant isoforms exist for other major VGICs \citep{abriel2010cardiac,benitah2010type}. In the mouse ventricle, repolarization is roughly delineated as two kinetic components: the rapidly inactivating current I\textsubscript{Kto}, conducted through K\textsubscript{v}4.2, and the slowly inactivating currents such as I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, conducted through K\textsubscript{v}1.5 and K\textsubscript{v}2.1 respectively. Although it is highly desirable to study individual K\textsubscript{v} isoforms to understand pathological mechanisms in the cardiac conduction system, only the collective activities as a whole can be measured in the form of currents using whole-cell voltage-clamp K\textsubscript{v} currents (I\textsubscript{K}) because of their slightly overlapping voltage-dependence of gating \citep{brouillette2004functional}. A data assimilation method, combining experimental data with mathematical/computational models, can used to estimate information on K\textsubscript{v} isoforms. Most existing studies separate individual K\textsuperscript{+} currents mathematically from I\textsubscript{K} by a curve-fitting decomposition procedure, assuming a curve of I\textsubscript{K} consists of the summation of exponential decay functions and a constant term; each represents the shape of K\textsuperscript{+} currents \citep{brunet2004heterogeneous}. However, this traditional approach is limited in its ability to scrutinize the kinetics and dynamics of K\textsubscript{v} in two-fold: 1) the curve-fitting I\textsubscript{K} decomposition only relies on the shape of K\textsuperscript{+} current without considering underlying kinetics and interactions of K\textsubscript{v} isoforms, and 2) it handles I\textsubscript{K} traces from the same cell with different conditions (i.e., clamp voltages) one-by-one individually. These limitations not only hinder \textit{in-silico} modeling results but, in some cases, make it impossible if modeling relies on kinetic data estimated from experimental data \citep{kim2022simulation}. There is an urgent need to develop new data assimilation methods that simultaneously delineate the underlying kinetics of multiple I\textsubscript{K} traces for a better understanding of the ion channels to provide insights into clinical applications, such as drug design.

Therefore, this paper proposes a new approach of \textit{concurrent data assimilation} for investigating the kinetics of K\textsubscript{v} isoforms from multiple I\textsubscript{K} traces measured in the same cell. A schematic flowchart of the proposed framework is shown in \textbf{Figure~\ref{fig:flow_chart}B} and compared with the current approach in \textbf{Figure~\ref{fig:flow_chart}A}. First, \textit{in-silico} models of K\textsubscript{v} and their currents are designed with parameters that control the kinetic rates and, in turn, determine the currents. It is worth mentioning that K\textsuperscript{+} currents are generated based on the mathematically simulated gating mechanism as opposed to the curve-fitting procedure, which only mimics the shape of the currents. Second, we perform a sensitivity analysis of the kinetic parameters using fractional factorial designs to identify the parameters that have significant impacts on the current generation. Several markers are defined that best represent the characteristics of the currents in various perspectives. Third, \textit{in-vitro} data of I\textsubscript{K} recordings are coupled with the proposed models by calibrating the parameters selected by the sensitivity analysis. We apply the proposed framework to glyco-altered IK recordings from our previous studies. Our data showed that reduced N-glycosylation in mouse cardiomyocytes is sufficient to cause dilated cardiomyopathy (DCM), which is the third most common cause of heart failure and the major reason for heart transplantation \citep{weintraub2017dilated}. Last, low-dimensional embedding is utilized to visually inspect the differences of the calibrated parameters in a collective way across the cells. Experimental results show that the proposed method provides reliable estimations of K\textsubscript{v} isoform kinetics of both the wild type and disease-altered groups and successfully captures cellular variability.
\begin{figure}
    \centering
    \includegraphics{figs/flow_chart.pdf}
    \caption{Flowcharts of (A) the current approach and (B) the proposed method.}
    \label{fig:flow_chart}
\end{figure}

The main advantage of this approach is that it analyzes multiple I\textsubscript{K} recordings concurrently based on the biophysical computer models, providing strong potential for modeling the kinetics and gating mechanism of the cardiac ion channels for studying heart diseases. Our contributions are summarized as follows:
\begin{itemize}
    \item We modeled the underlying kinetics of K\textsubscript{v} isoforms from whole-cell I\textsubscript{K} recordings based on biophysical principles rather than curve-fitting procedures.
    \item We developed a concurrent data assimilation routine that harnesses multiple protocols simultaneously rather than a single protocol independently one-by-one, resulting in comprehensive cellular-level modeling.
    \item The proposed approach generates separate K\textsubscript{v} models and visualize the distributions of the kinetic parameters for each cell, which allows analyzing the cellular variability. 
    \item Our method captures differences in channel kinetics given the disease-related perturbation, such as reduced glycosylation.
\end{itemize}
The proposed data assimilation framework has strong potential to serve as a novel complementary data analytical tool to \textit{in-vitro} experiments.

@@

\subsection{Concurrent Assimilation of Functional Data}
Data assimilation is a systematic procedure to find the optimal configuration and state of computational/mathematical models by coupling it with experimental data. Experimental data $\mathcal{D}$ are observations of a real process $\mathcal{R}$ that represents scientific phenomena under investigation. The output of physical experiments $y^{\mathcal{D}}(x)$, given input $x$, inevitably contains errors for various reasons, such as noise in measurement or experimental environment. Suppose $\mathcal{D}$ and $\mathcal{R}$ can be related as follows in Equation~\ref{eq:rd_relation}, where $\epsilon$ is the error term. 
\begin{equation}
    \label{eq:rd_relation}
    y^{\mathcal{R}}(x) = y^{\mathcal{D}}(x) + \epsilon
\end{equation}
Let $y^{\mathcal{M}}(x|\theta)$ denote the output from a computer model $\mathcal{M}$, given parameters $\theta$. Assume that there are discrepancies $\delta(x|\theta)$ for the current states of parameters as follows in Equation~\ref{eq:dm_relation}. 
\begin{align}
    \label{eq:dm_relation}
    y^{\mathcal{D}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) \text{, so} \\
    y^{\mathcal{R}}(x) &= y^{\mathcal{M}}(x|\theta) + \delta(x|\theta) + \epsilon
\end{align}
Our goal in data assimilation is to calibrate $\theta$ to find the best model states that minimize $\delta(x|\theta)$, while satisfying biophysical constraints. By doing that, \textit{in-silico} models $\mathcal{M}$ are coupled with \textit{in-vitro} experimental data $\mathcal{M}$, which provides two complementary angles to study the real process $\mathcal{R}$.

Traditionally, bi-/tri-exponential functions as described in Equation~\ref{eq:expl_fitting} have served as $y^{\mathcal{M}}(x|\theta)$, where $\theta=\{\mathrm{A}_{i}(x),\tau_{i}(x)\}$. Note that the model parameters $\mathrm{A}_{i}(x)$ (amplitude) and $\tau_{i}(x)$ (time constant) depend on input data $x$, which are clamp voltages for whole-cell I\textsubscript{K} recordings. In general, multiple input protocols are applied to a myocyte producing a set of I\textsubscript{K} recordings to study the voltage-dependent characteristics of VGICs. However, these exponential functions cannot analyze simultaneously, leading to potential information loss in the cellular-level dynamics of ion channels. In contrast, computer models of VGICs can generate currents given one set of parameters because they are designed by biophysical principles encoding voltage dependence. In this study, $\delta$ is defined by the summation of root-mean-square errors (RMSEs) as given in Equation~\ref{eq:rmse}. RMSE evaluates the goodness-of-fit between a I\textsubscript{K} recording recording and model prediction from the end of a holding potential $t_v^{(h)}$ to the end of voltage step $t_v^{(e)}$. RMSEs from the same cell are summed up to guide the optimization procedure calibrating the models in a concurrent way for clamp voltage $v=1,2,\cdots,n$.
\begin{equation}
    \label{eq:rmse}
    \delta = \sum_{v=1}^{n} \sqrt{\int_{t_v^{(h)}}^{t_v^{(e)}}\frac{(y_v^{\mathcal{D}}(t) - y_v^{\mathcal{M}}(t|\theta))^2}{t_v^{(e)}-t_v^{(h)}}dt}
\end{equation}

This approach differs from previous studies in two-fold. First, unlike the traditional curve fitting using the exponential model described in Equation~\ref{eq:expl_fitting}, the suggested method includes multiple protocols simultaneously for comprehensive cellular-level modeling. Second, it calibrates computer models directly to I\textsubscript{Ksum} recordings, while the previous studies use statistics estimated from the data \citep{du2015statistical, du2017silico, kim2022simulation}. Calibrating to the current traces themselves raises a challenge in optimization but has advantages when it is hard to estimate statistics from data reliably \citep{kim2022simulation}. We developed the box-constrained nonlinear optimization routine using nonlinear optimization algorithms with multi-random initial points to minimize $\delta$. Box constraints mean that $\theta$ has a lower and upper bound for each dimension, so solution space is constrained in a hypercube. In this way, the optimization loop can be controlled by users, allowing them to blend their domain knowledge into the modeling. Note that it is possible because the suggested method is based on biophysical models that provide interpretability of $\theta$ in channel kinetics rather than the curve fitting. Global optimum is not guaranteed due to the complexity of our models. We developed the multi-random-start scheme to avoid local optima and find the solution as close to the optimum as possible. Latin hypercube designs and parallel computations are used to sample initial points and run them on multicores to compensate for the increased computational burden. This work is implemented in MATLAB R2022a, and the code is available as online supplemental material.

@@

\subsection{Kinetics Modeling} \label{s:results.kinetics}
The \textit{in-silico} modeling predicts that reduced N-glycosylations cause significant changes in channel kinetics and gating behavior. \textbf{Figure~\ref{fig:kinetics_ikto}} presents the \textit{in-silico} modeling results showing kinetics variables to voltage relationships of K\textsubscript{v}4.2. There is a very slight rightward shift in steady-state activation (SSA) values along the voltage axis (\textbf{Figure~\ref{fig:kinetics_ikto}A}), and steady-state inactivation (SSI) is reduced up to 0.25 at voltages lower -10 mV (see \textbf{Figure~\ref{fig:kinetics_ikto}B}). Activation and inactivation time constants ($\tau_{a}^{(1)}$ and $\tau_{i}^{(1)}$) to voltage relationships are shown in \textbf{Figure~\ref{fig:kinetics_ikto}C} and \textbf{D}. $\tau_{a}^{(1)}$ for MGAT1KO is greater than WT in range of $>$ -20 mV. $\tau_{i}^{(1)}$ for MGAT1KO shifted toward left. The shifts in SSA and SSI likely cause the reductions in I\textsubscript{Kto} magnitudes, and the increased time constants explain delays in the decaying phase observed in the \textit{in-vitro} experiments.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kinetics_ikto.pdf}
    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kto} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, and (D) inactivation time constant.}
    \label{fig:kinetics_ikto}
\end{figure}

\textbf{Figure~\ref{fig:kinetics_rec}} shows kinetics modeling results of the channels producing the delayed rectifier currents, I\textsubscript{Kslow1} and I\textsubscript{Kslow2}, and steady-state current I\textsubscript{Kss}. The most prominent result is that time constants of activation and inactivation in I\textsubscript{Kslow1} ($\tau_{a}^{(2)}$ and $\tau_{i}^{(2)}$) are estimated to be significantly slow in MGAT1KO, i.e., $\sim$ 3.5 and 348.9 ms, respectively (see \textbf{Figure~\ref{fig:kinetics_rec}C} and \textbf{D}). Similar to I\textsubscript{Kto}, there is a small rightward shift in SSA at voltaged $>$ -10 mV for MGAT1KO (see \textbf{Figure~\ref{fig:kinetics_rec}A}). SSI is also shifted $\sim$ 10 mV to the right (see \textbf{Figure~\ref{fig:kinetics_rec}B}). The slow time constants and increased SSI are likely to address the flattened and reduced currents of MGAT1KO. Time constant inactivation of I\textsubscript{Kslow2} ($\tau_{i}^{(3)}$) is also increased (see \textbf{Figure~\ref{fig:kinetics_rec}E}), but activation of I\textsubscript{Kss} ($\tau_{a}^{(4)}$) do not show significant differences between the two groups (see \textbf{Figure~\ref{fig:kinetics_rec}F}). It is worth mentioning that the proposed method extrapolates $\tau_{i}^{(2)}$ that is greater than 5 s from the 4.5-s-pulse protocol.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/kinetics_rec.pdf}
    \caption{Kinetics functions and their relationships to voltage of I\textsubscript{Kslow1} (A) steady-state activation, (B) steady-state inactivation, (C) activation time constant, (D) inactivation time constant, (E) I\textsubscript{Kslow2} inactivation time constant, and (F) I\textsubscript{Kss} activation time constant.}
    \label{fig:kinetics_rec}
\end{figure}

We aggregated calibration results into one to analyze the parameters in a collective way. It resulted in a $61\times20$ matrix (i.e., 31 observations in WT and 30 in MGAT1KO, and there are 20 calibration parameters). Then we applied t-distributed stochastic neighbor (t-SNE) embedding to encapsulate the high-dimensional data into three-dimensional space so that we could visualize each data point. \textbf{Figure~\ref{fig:embd_result}} presents the visualization of the t-SNE embedding of the calibration results. It shows not only clear differences between the two groups but also variances of cells (i.e., data points) within the groups and across the entire dataset.
\begin{figure}[!ht]
    \centering
    \includegraphics{figs/embd_result.pdf}
    \caption{t-SNE embedding of calibration parameters.}
    \label{fig:embd_result}
\end{figure}
